Manuscript for foodborne-outbreak Project

### 2023-02-24 code/analysis_code/exploratory_analysis.qmd
part 2 continued work

# Food Type and Hospitalizations
ggplot(training_data_final %>% 
         filter(`IFSAC Category` != "Multiple") %>% 
         filter(`IFSAC Category` != "Other"), 
       aes(x = `IFSAC Category`, y = Hospitalizations)) + 
  geom_col() + 
  labs(x = "Food Type", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'")
ggsave("Food_type_v_hospitalization.png", path = "results")

# Locations and Hospitalizations
Location_v_hospitalization=ggplot(training_data_final, aes(x = Location, y = Hospitalizations)) +
  geom_col() + 
  labs(x = "Locations", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'")+
  scale_x_discrete(guide=guide_axis(n.dodge=3))
Location_v_hospitalization

ggsave("Location_v_hospitalization.png", path = "results")

@@

#save graph
ggsave("bacterial_etiology_v_outbreak_frequency.png", path = "results")

@@

#save graph
ggsave("food_type_v_illness_frequency.png", path = "results")

@@

#save graph
ggsave("Location_v_Illness.png", path = "results")

>>>>

# Food Type and Hospitalizations
ggplot(training_data_final %>% 
         filter(`IFSAC Category` != "Multiple") %>% 
         filter(`IFSAC Category` != "Other"), 
       aes(x = `IFSAC Category`, y = Hospitalizations)) + 
  geom_col() + 
  labs(x = "Food Type", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'")
ggsave("Food_type_v_hospitalization.png", path = image_path)

# Locations and Hospitalizations
Location_v_Hospitalization <- ggplot(training_data_final, aes(x = Location, y = Hospitalizations)) +
  geom_col() + 
  labs(x = "Locations", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'")+
  scale_x_discrete(guide=guide_axis(n.dodge=3))

Location_v_Hospitalization

ggsave("Location_v_hospitalization.png", path = image_path)

@@

# Save Graph
ggsave("bacterial_etiology_v_outbreak_frequency.png", path = image_path)

@@

# Save Graph
ggsave("food_type_v_illness_frequency.png", path = image_path)

@@

# Save Graph
ggsave("Location_v_Illness.png", path = image_path)

### 2023-02-24 products/manuscript/Manuscript.qmd
part 2 continued work

We adopted two datasets for this study.
The first dataset was uploaded onto Kaggle in 2017 and contains data from the Centers for Disease Control and Prevention (CDC) on foodborne disease outbreaks from 1998 to 2015 @centersfordiseasecontrolandprevention2017.
There are 12 variables and 16,950+ observations.

The second dataset was collected from National Outbreak Reporting System (NORS), Centers for Disease Control and Prevention (CDC) website.
This dataset covers the reports of foodborne and waterborne disease outbreaks and enteric (intestinal) disease outbreaks spread by contact with environmental sources, infected people or animals, and other means from 2009 to 2020.
The raw datasets may be viewed within the repository of this project (data/raw_data/).

@@

We have several questions we would like our datasets to answer: (1) what are the most frequent locations that foodborne disease outbreaks occur (2) what pathogen species are most frequently associated with foodborne outbreaks, and (3) what type of food is most commonly involved in foodborne outbreaks. The outcomes of interest would therefore be location (for whether more outbreaks occur in restaurants/homes), species of pathogen, and food type most often associated with outbreaks.

>>>>

We adopted two datasets for this study.
The first dataset [@cdc2020]was uploaded onto Kaggle in 2017 and contains data from the Centers for Disease Control and Prevention (CDC) on foodborne disease outbreaks from 1998 to 2015 [@cdc2017].
There are 12 variables and 16,950+ observations.

The second dataset was collected from the National Outbreak Reporting System (NORS), Centers for Disease Control and Prevention (CDC) website .
This dataset covers the reports of foodborne and waterborne disease outbreaks and enteric (intestinal) disease outbreaks spread by contact with environmental sources, infected people or animals, and other means from 2009 to 2020.
The raw datasets may be viewed within the repository of this project (data/raw_data/).

@@

We have several questions we would like our datasets to answer: (1) what are the most frequent locations that foodborne disease outbreaks occur, (2) what pathogen species are most frequently associated with foodborne outbreaks, and (3) what type of food is most commonly involved in foodborne outbreaks.
The outcomes of interest would therefore be location (for whether more outbreaks occur in restaurants/homes), species of pathogen, and food type most often associated with outbreaks.
